---
title: "Ch_07_Regression Analysis"
author: "Murtaza Haider"
date: "January 16, 2016"
output: html_document
---



##Download data

**Downloading** data from the website. The download command downloads data in the working directory.

Data are downloaded, loaded, and names of fields are read from the file:

```{r}
setInternet2(TRUE) # solution for https files

download.file("https://sites.google.com/site/econometriks/docs/housing88.RData", "housing88.RData")

load("housing88.RData")
setwd("C:/Users/Regionomics/Documents/Research_2/All_else_being_equal/Chapters/Ch07_Regression/Data/Regression_housing/Housing prices/88_units")

names(housing88)
```


**Loading** required packages and **attaching** the data set.
```{r}
attach(housing88)
```


```{r}
library(lattice)
lattice.options(default.theme = standard.theme(color = FALSE))
trellis.device(color = FALSE)
```


### Figure 7.3

```{r,results='asis'}
options(scipen=10)
xyplot(hprice ~ sqrft, data = housing88,
       type = c("p", "g", "r"), pch=16, lwd=2, col.symbol = "black",
       xlab = "size of house in square feet",
       ylab= "housing price")
```


### Table 7.5

```{r,results='asis'}
library(stargazer)
stargazer(housing88, type="html", digits=2)
```

### Figure 7.4

```{r, results= 'asis'}

histogram(lotsize, nint=10,
          xlab="size of lot in square feet", col=c("dark grey"))

```

### Figure 7.5

```{r}

mod1<-(lm(hprice~bdrms+lotsize+sqrft+colonial))
summary(mod1)

```

### Table 7.6

```{r, results='asis'}

stargazer(mod1,type='html', no.space=TRUE, align=TRUE,
          dep.var.labels=c("House price"),
          covariate.labels=c("number of bedrooms", 
                             "size of lot in square feet",
                             "size of house in square feet",
                             "=1, if Colonial style"))

```


### Table 7.7


```{r}
library(psych)
library(xtable)
library(htmlTable)
```


```{r}
f <- function(x) c(obs=length(x), mean=mean(x), sd=sd(x))
tres<-as.matrix(aggregate(hprice ~ bdrms, FUN=c("f")))
colnames(tres)[2:4]<-c("n", "mean", "std. dev.")
htmlTable(txtRound(tres,0))
```


### Table 7.8

```{r, results='asis'}
mod2<-lm(hprice ~ factor(bdrms))
stargazer(mod2,type='html', no.space=TRUE, align=TRUE, digits=2,
          dep.var.labels=c("House price"))
```


### Table 7.9

```{r, results='asis'}
mod3<-lm(hprice ~ bdrms)
mod4<-lm(hprice ~ bdrms + lotsize )
mod5<-lm(hprice ~ bdrms + sqrft )
mod6<-lm(hprice ~ bdrms + lotsize + sqrft )
mod7<-lm(hprice ~ bdrms + lotsize + sqrft + colonial)

stargazer(mod3, mod4, mod5, mod6, mod7, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("house price"),
           covariate.labels=c("number of bedrooms", 
                             "size of lot in square feet",
                             "size of house in square feet",
                             "=1, if Colonial style"))

```

### Table 7.11
names(housing88)

```{r, results='asis'}
mod3<-lm(lprice ~ lsqrft)
mod4<-lm(lprice ~ colonial)
mod5<-lm(lprice ~ bdrms + bedsq )

stargazer(mod3, mod4, mod5, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("log of house price"),
           covariate.labels=c("log of built area", 
                              "=1, if Colonial style",
                             "number of bedrooms",
                             "beds squared"))

```

# Teaching Ratings

```{r}
setInternet2(TRUE) # solution for https files

download.file("https://sites.google.com/site/statsr4us/intro/software/rcmdr-1/TeachingRatings.rda", "TeachingRatings.rda")

load("TeachingRatings.rda")
names(TeachingRatings)
attach(TeachingRatings)
```

### Table 7.13


```{r, results='asis'}

mod3<-lm(eval ~ gender )
mod4<-lm(eval ~ gender + minority + native)
mod5<-lm(eval ~ gender + minority + native + tenure + division +credits)
mod6<-lm(eval ~ gender + minority + native + tenure + division +credits + 
           beauty)

stargazer(mod3, mod4, mod5, mod6, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("teaching evaluation score"),
           covariate.labels=c("female", "minority instructor",
                              "NON-native English speaker", "tenured professor",
                              "lower division", "single credit", 
                              "normalized score of instructor's appearance",
                              "intercept"))

```


### Table 7.14

```{r, echo=FALSE}
TeachingRatings$weights <- students/allstudents
attach(TeachingRatings)
library(sandwich)
library(lmtest)
library(multiwayvcov)
```



```{r, results='asis'}

modw1<-lm(eval ~ gender + minority + native + tenure + division +credits + beauty)
modw2<-lm(eval ~ gender + minority + native + tenure + division +credits +
           beauty, weight=weights)

vcov_prof <- cluster.vcov(modw2, TeachingRatings$prof)
# coeftest(modw2, vcov_prof)

robust.se <- sqrt(diag(vcov_prof))

stargazer(modw1, modw2, modw2, type="html", align=TRUE, no.space=TRUE, 
           se=list(NULL, NULL, robust.se),
          digits=3, dep.var.labels=c("teaching evaluation score"),
          column.labels=c("un-weighted model","weighted model",
                         "weighted model w/ clustered std. errors"),
          covariate.labels=c("female", "minority instructor",
                             "NON-native English speaker", "tenured professor",
                              "lower division", "single credit", 
                              "normalized beauty score",
                              "intercept"))

```



# Expenses

```{r, echo=FALSE}
setInternet2(TRUE) # solution for https files

download.file("https://sites.google.com/site/econometriks/docs/hhld_expenses.rda", "hhld_expenses.rda")

load("hhld_expenses.rda")
names(Dataset)
attach(Dataset)
```

## Alcohol

### Table 7.16

```{r,results='asis'}
stargazer(Dataset, type="html", digits=2)
```


### Table 7.17
```{r}
f <- function(x) c( mean=mean(x), obs=length(x))
tres<-as.matrix(aggregate(alcoh ~ kids.f, FUN=c("f")))
colnames(tres)[1:3]<-c("children in hhld", "mean", "n")
htmlTable(txtRound(tres,1))
```

### Table 7.18

```{r, results='asis'}
lm.1 <- lm(alcoh ~ adults, data=Dataset)
lm.2 <- lm(alcoh ~ adults + income, data=Dataset)
lm.3 <- lm(alcoh ~ adults + income + kids, data=Dataset)
lm.4 <- lm(alcoh ~ adults + income + kids, subset= kids>0 ,
        data=Dataset)

stargazer(lm.1, lm.2, lm.3, lm.4, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("spending on alcohol"),
          column.labels=c("alcohol", "alcohol, households with children"),
          column.separate = c(3,1),
           covariate.labels=c("adults in hhld", "weekly income",
                              "children in hhld"))
```
tres


### Figure 7.7

```{r, results='asis'}
barplot(tapply(alcoh,kids,mean), ylim =c(0,40),
        ylab="Weekly household spending on alcohol ($)", 
        xlab = "No. of children",
        main="Weekly spending on alcohol (Canadian $)")
```

### Table 7.19

```{r, results='asis'}
lm.3 <- lm(alcoh ~ income + factor(adults) +  factor(kids), data=Dataset)
stargazer(lm.3, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("spending on alcohol"))

```
  
  
## Food


### Table 7.20
```{r}
f <- function(x) c( mean=mean(x), obs=length(x))
tres<-as.matrix(aggregate(food ~ kids.f, FUN=c("f")))
colnames(tres)[1:3]<-c("children in hhld", "mean", "n")
htmlTable(txtRound(tres,1))
```  

### Table 7.21

```{r, results='asis'}
lm.1 <- lm(food ~ adults, data=Dataset)
lm.2 <- lm(food ~ adults + income, data=Dataset)
lm.3 <- lm(food ~ adults + income + kids, data=Dataset)
lm.4 <- lm(food ~ adults + income + kids, subset= kids>0 ,
        data=Dataset)

stargazer(lm.1, lm.2, lm.3, lm.4, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("spending on food"),
          column.labels=c("food", "food, households with children"),
          column.separate = c(3,1),
           covariate.labels=c("adults in hhld", "weekly income",
                              "children in hhld"))

```



### Table 7.22

```{r, results='asis'}
lm.3 <- lm(food ~ income + factor(adults) +  factor(kids), data=Dataset)
stargazer(lm.3, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("spending on food"))

```

### Table 7.23

```{r, results='asis'}
mod1<-lm(hprice~ lotsize+sqrft+bdrms)

res1 <- resid(mod1)
res2<-res1^2

mod2 <- lm(res2 ~ lotsize+sqrft+bdrms)

stargazer(mod1, mod2, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("housing price", "residuals squared"),
          column.labels=c("BP Test, original model",
          "BP Test, Residual Squared model"),
          covariate.labels=c("size of lot in sqr. ft.", "house size in sqr. ft.",
                              "number of bedrooms"))
```

### Table 7.24

```{r, results='asis'}
mod1<-lm(lprice~ llotsize+lsqrft+lnrooms)

res1 <- resid(mod1)
res2<-res1^2

mod2 <- lm(res2 ~ llotsize+lsqrft+lnrooms)

stargazer(mod1, mod2, type="html", align=TRUE, no.space=TRUE, 
          digits=2, dep.var.labels=c("log of housing price", "residuals squared"),
          column.labels=c("BP Test, original model",
          "BP Test, Residual Squared model"),
          covariate.labels=c("size of lot in sqr. ft.", "house size in sqr. ft.",
                              "number of bedrooms"))
```


### Table 7.25

```{r, results='asis'}
mod1<-lm(hprice~ lotsize+sqrft+bdrms)

library(sandwich)
cov <- vcovHC(mod1, type = "HC1")
robust.se <- sqrt(diag(cov))

stargazer(mod1, mod1, type="html", align=TRUE, no.space=TRUE,
          se=list(NULL, robust.se),
          digits=2, column.labels=c("Default SE", "Robust SE"),
          covariate.labels=c("size of lot in sqr. ft.", "house size in sqr. ft.",
                              "number of bedrooms"))
```


### Table 7.26

```{r, results='asis'}
mod1<-lm(hprice~ bdrms)
mod2<-lm(hprice~ bdrms +sqrft)
mod3<-lm(hprice~ bdrms +sqrft+lotsize)


stargazer(mod1, mod2, mod3, type="html", align=TRUE, no.space=TRUE,
          digits=2, column.labels=c("beds", "beds+sqrft", "beds+sqrft+lotsize"),
          covariate.labels=c("number of bedrooms",
                             "house size in sqr. ft.", "size of lot in sqr. ft."))
```
      
      
      
      
---