---
title: "Chapter 8 Binary Logit"
author: "Murtaza Haider"
date: "January 24, 2016"
output: html_document
---

###load data

Data are downloaded, loaded, and names of fields are read from the file:

```{r}
setInternet2(TRUE) # solution for https files
download.file("https://sites.google.com/site/econometriks/docs/R_smoking.rda", "R_smoking.rda")
load("R_smoking.rda")

names(smoking)
```


**Loading** required packages and **attaching** the data set.

```{r}
library(xtable)
library(psych)
attach(smoking)
library(htmlTable)
library(stargazer)
```

####Descriptive stats on continuous varaiables

**Table 8.4**

```{r, results='asis'}
t1<-table(smoker)
tab <- xtable(cbind(t1,'t1(%)'=t1/sum(t1)*100))
htmlTable(txtRound(tab,2))

```
***

**Table 8.5**
```{r, results='asis'}

smoking$smoker2<-factor(smoking$smoker)

f1 <- function(x) c( mean=mean(x), sd= sd(x), obs=length(x))

tab<-rbind(education=f1(educ),'age in years'=f1(age),
      'household income'=f1(income), 
       'cigarette price in cents'=f1(pcigs79),
      'smoker' =f1(as.numeric(smoking$smoker2)-1))
htmlTable(txtRound(tab,2))

```

...
...

**Descriptive statistics by smoker status**


**Table 8.6**
```{r, results='asis'}
tab <- xtable(describeBy(cbind(educ, age, income, pcigs79), smoker, mat=T, skew=F, ranges=F))
names(tab)[2] <-c( "smoker")
print(tab, type="html")
```

***

###Other versions of Table 8.6

```{r, results='asis'}
t1<-(aggregate(cbind(educ, age, income, pcigs79) ~ smoker, data=smoking,f1))
htmlTable(txtRound(t(t1),2))


t1<-aggregate(cbind(income,age, educ, pcigs79) ~ smoker, data=smoking, f1)
tab<-xtable( as.matrix(t(t1)) )
print(tab, type="html")
```


```{r, results='asis'}
t1<-aggregate(cbind(income,age, educ, pcigs79) ~ smoker+1, data=smoking, f1)
t2<-aggregate(cbind(income,age, educ, pcigs79) ~ 1, data=smoking, f1);

t3<-as.matrix(t(t1))
t4<-as.matrix(rbind("total",t(t2)))
t5<-cbind(t3,t4)

tab<-xtable(t5)
print(tab, type="html")
htmlTable(txtRound(t5,0,excl.rows=c(1:3,5,6,8,9,11,12)))

```
***

##OLS model##

**Table 8.7**

```{r, results='asis'}
ols.1<- lm((as.numeric(smoker)-1) ~ age +educ + pcigs79 +income,
           data = smoking)
ols.1a <- xtable(summary(ols.1))
print(ols.1a, type="html") 

ols.2<- lm((as.numeric(smoker)-1) ~ age +educ + pcigs79 +inc.10k)
ols.2a <- xtable(summary(ols.2))
print(ols.2a, type="html") 

## Alternate version

stargazer(ols.1,ols.2, type="html", align=TRUE, no.space=TRUE, 
          digits=4, dep.var.labels=c("smoker"),
          column.labels=c("income", "income/10k"),
          covariate.labels=c("age in years", 
                             "education in years",
                             "price of a pack", "household income",
                             "income in 10k"))

```


***


##Logit models##

**Table 8.9**

```{r, results='asis'}
GLM.1 <- glm(smoker ~ age +educ + pcigs79 + income, family=binomial(logit))
coef.vector1 <-exp(GLM.1$coef)

stargazer(GLM.1, type="html", align=TRUE, no.space=TRUE, 
          digits=4, dep.var.labels=c("smoker"),
          column.labels=c("income", "income/10k"),
         covariate.labels=c("age in years", "education in years",
                           "price of a pack", "household income"))
```

...


**Exponentiated coefficients**
**Table 8.10**

```{r, results='asis'}
GLM.1 <- glm(smoker ~ age +educ + pcigs79 + income, family=binomial(logit))
coef.vector1 <-exp(GLM.1$coef)

GLM.2 <- glm(smoker ~ age +educ + pcigs79 + inc.10k, family=binomial(logit))
coef.vector2 <-exp(GLM.2$coef)

stargazer(GLM.1, GLM.2, coef= list(coef.vector1, coef.vector2), 
           digits=4, dep.var.labels=c("smoker"),
          column.labels=c("income", "income/10k"), 
          type="html",
          covariate.labels=c("age in years", "education in years",
                           "price of a pack", "household income",
                           "income in 10k"))
```



** **

##Probit models##


**Figure 8.9**

```{r}
GLM.p<-(glm(smoker ~ age +educ + pcigs79 + inc.10k, data=smoking, family=binomial(probit)))
summary(GLM.p)
```

***

**Table 8.12**


```{r, results='asis'}

stargazer(ols.2, GLM.2,GLM.p, digits=4, dep.var.labels=c("smoker"),
          type="html",
          covariate.labels=c("age in years", 
                             "education in years",
                             "price of a pack", "income in 10k"))
```

#Zelig 

## Figure 8.15, 8.16 and more

```{r,message=FALSE,echo=FALSE}
library(Zelig)
```


```{r, eval=T}
z.out1 <- zelig(smoker ~ age + educ  + pcigs79 + income,
                model = "logit", data = smoking)
x.out1 <- setx(z.out1, age = 41.88414,  educ=12.19118, 
               income = 19359.18, pcigs79 = 60.97193)
s.out1 <- sim(z.out1, x = x.out1, num=100000)
summary(s.out1)

x.out2 <- setx(z.out1, age = 41.88414,  educ=12.19118, 
               income = 19359.18, pcigs79 = 76.2149125)
s.out2 <- sim(z.out1, x = x.out2, x1 = x.out1, num=100000)
summary(s.out2)
```

```{r,fig.height=7, eval=T}
plot(s.out1)
plot(s.out2)
```


###Figures 8.16, 8.17 and 8.18

```{r}
library (erer)
ma <- glm(smoker ~ age + educ + inc.10k + pcigs79, x = TRUE,
data = smoking, family = binomial(link = "probit"))
ea <- maBina(w = ma, x.mean = TRUE, rev.dum = TRUE)
summary(ma)
ea

edu.1 <- maTrend(q = ea, nam.c="educ")
plot(edu.1)
```

###Grouped Logit###

```{r}

download.file("https://sites.google.com/site/econometriks/docs/ny_grouped.rda", "ny_grouped.rda")
load("ny_grouped.rda")

attach(ny_grouped)
```



**Table 8.15**

```{r, results='asis'}
f1 <- function(x) c( mean=mean(x), sd= sd(x), obs=length(x))

tab<-rbind("average autos/hhld"=f1(hh_autos),
           'average persons/hhld'=f1(hh_size),
      'distance from CBD'=f1(d_cbd), 
       'total number of trips'=f1(n_trips),
      'total transit trips' =f1(n_transit))
htmlTable(txtRound(tab,2))

```
...

**Figure 8.20 and Table 8.17**

```{r, results='asis'}
ny_grouped$non_transit <- with(ny_grouped, n_trips-n_transit)
detach(ny_grouped)
attach(ny_grouped)

mod1 <- glm(cbind( n_transit, non_transit) ~ hh_autos + hh_size + 
            d_cbd, data = ny_grouped,
            family = binomial(link="logit"),trace=T)

stargazer(mod1, type="html",
          covariate.labels=c("avg. autos/hhld", "avg. persons/ hhld",
                             "distance from cbd")) 
```

**Figure 8.19**

```{r}

res<-(data.frame("observed transit trips"=(ny_grouped$n_transit), "forecasted transit trips"=fitted(mod1)*ny_grouped$n_trips))

head(res)
plot(res)
```



