---
title: "Beauty, Chapter 4"
author: "Murtaza Haider"
date: "August 27, 2014, revised Jan. 05, 2016"
output: html_document
---


##Download data

**Downloading** data from the website. The download command downloads data in the working directory.

Data are downloaded, loaded, and names of fields are read from the file:

```{r}
setInternet2(TRUE) # solution for https files

download.file("https://sites.google.com/site/statsr4us/intro/software/rcmdr-1/TeachingRatings.rda", "TeachingRatings.rda")

load("TeachingRatings.rda")

setwd("C:/Users/Murtaza/Google Drive/AEBE/Data/ch.04/R_ch04")

names(TeachingRatings)
```

**Loading** required packages and **attaching** the data set.

```{r}
library(xtable)
library(psych)
library(htmlTable)
attach(TeachingRatings)
```

####Descriptive stats on continuous varaiables

***TAble 4-18***

```{r, results='asis'}

tab <- xtable(describe(cbind(age, beauty, eval, students, allstudents), 
                       skew=F, ranges=T))
print(tab, type="html")

```
** **

**Subset data by professors**

Professor specific variables:

- age
- beauty


***Identifying duplicates for subsetting***


```{r}
profs <- subset(TeachingRatings, !duplicated(TeachingRatings$prof))
```


```{r, warning=FALSE, message=FALSE}
attach(profs)
```

**Descriptive statistics**


***Table 4-19***
```{r, results='asis'}
tab <- xtable(describe(cbind(age, beauty), skew=F, ranges=F))
print(tab, type="html")

```

****

**Evaluation Factors:**


- minority 
- Gender 
- Native 
- Tenure 


```{r, warning=FALSE, message=FALSE}
attach(TeachingRatings)
```

#### Average evaluation by gender 


**psych package**

***Table 4-20***

```{r, results='asis'}
tab <- xtable(describeBy(eval, gender, mat=T, skew=F, ranges=T))
print(tab, type="html")

```
****



```{r xtable1, results="asis"}
tab<-xtable(cbind("Teaching Evaluations"=tapply(eval, gender, mean)))
print(tab, type="html")
```


#### Average evaluation by tenure 

***Table 4-21***

```{r}
tab<-cbind("Teaching Evaluations"=tapply(eval, tenure, mean))
htmlTable(txtRound(tab,2),rowlabel = "Tenure status")

```


```{r xtable2, results="asis"}
tab<-xtable(cbind("Teaching Evaluations"=tapply(eval, tenure, mean)))
print(tab, type="html")
```

#### Average evaluation by tenure and gender

***Table 4-22***

```{r}
tab<-cbind("Teaching Evaluation"=tapply(eval, list(tenure, gender), mean))

htmlTable(txtRound(tab,2), 
          cgroup = c("Gender"), n.cgroup = c(2),
          rowlabel = "Tenure Status")

```



****
***Table 4-23 and 4-24***

```{r}
f <- function(x) c(obs=length(x), mean=mean(x), sd=sd(x))
tres<-as.matrix(aggregate(eval ~ gender+tenure, FUN=c("f")))
htmlTable(txtRound(tres,0,excl.cols = c(1,2,4,5)))

colnames(tres)[3:5]<-c("n", "mean", "std. dev.")
htmlTable(txtRound(tres,0,excl.cols = c(1,2,4,5)))
```


#### Average evaluation by tenure, gender, and minority

***Table 4-25***

```{r}
f <- function(x) c(obs=length(x), mean=mean(x), sd=sd(x))
tres<-as.matrix(aggregate(eval ~ gender+tenure+minority, FUN=c("f")))
colnames(tres)[4:6]<-c("n", "mean", "std. dev.")
htmlTable(txtRound(tres,0,excl.cols = c(1,2,3,5,6)))
```



**Summarising eval and beauty by gender**

***Table 4-26***
```{r}
f <- function(x) c(obs=length(x), mean=mean(x), sd=sd(x))
tres<-t(as.matrix(aggregate(cbind(eval, beauty) ~ gender, FUN=c("f"))))
rownames(tres)[2:7]<-c("n", "mean", "std. dev.", "n", "mean", "std. dev.")
htmlTable(txtRound(tres,0, excl.rows = c(3,4,6,7)),
          rgroup=c("","evaluation score","beauty score"),
          n.rgroup = c(1,3,3))
```


#### Average evaluation by minority

```{r xtable4, results="asis"}
tab<-xtable(cbind("Teaching Evaluation"=tapply(eval, minority, mean)))
print(tab, type="html")
```



****
#### Average evaluation by native English speaking

```{r xtable6, results="asis"}
tab<-xtable(cbind("Teaching Evaluation"=tapply(eval, native, mean)))
print(tab, type="html")
```
..





#### Correlation between beauty and evaluation

***Table 4-27***

```{r, results='asis' }

myvars<- c("eval", "beauty", "age", "students", "allstudents")
x <- TeachingRatings[myvars]
tab <-xtable(cor(x), digits=3)
print(tab, type="html")
```




****

##Reproducing Table 1 from Hamermesh paper


**Generate SD Function**
```{r}
my.wsd<-  function(w, x, wx, n)  sqrt(sum(w*(x-wx)^2)/((n-1)*sum(w)/n))
```



**Generate Percent evaluating**

```{r}
TeachingRatings$p.eval <- students/allstudents*100 
```

**All Results**

```{r, warning=FALSE, message=FALSE}
attach(TeachingRatings)


```

```{r}

wm.course <-weighted.mean(eval,students); wm.course
wsd.course<-my.wsd(students,eval, wm.course, length(eval)) ; wsd.course
m.studs <- mean(allstudents); sd.studs <-sd(allstudents); m.studs
m.peval <-mean(p.eval); m.peval

tabgen<-xtabs(students~gender)/sum(students); tabgen
tabmin<-xtabs(students~minority)/sum(students); tabmin
tabeng<-xtabs(students~native)/sum(students); tabeng
tabten<-xtabs(students~tenure)/sum(students); tabten
tabdiv<-table(division)/sum(table(division)); tabdiv
tabmin<-xtabs(students~minority)/sum(students); tabmin
courses <- length(eval); courses
faculty <- length(unique(prof)); faculty

all.res<- rbind(wm.course, wsd.course, m.studs, sd.studs, m.peval, female=tabgen[2],
      minority=tabmin[2], 'native english' =tabeng[2], tenure = tabten[2],
      "lower division" = tabdiv[2], courses, faculty); all.res

colnames(all.res) <- "all courses"; all.res
```

****


**Subset Lower**

```{r, warning=FALSE, message=FALSE}
lower <- subset(TeachingRatings, subset=division =="lower")
attach(lower)
```

```{r}

wm.course <-weighted.mean(eval,students); wm.course
wsd.course<-my.wsd(students,eval, wm.course, length(eval)) ; wsd.course
m.studs <- mean(allstudents); sd.studs <-sd(allstudents); m.studs
m.peval <-mean(p.eval); m.peval

tabgen<-xtabs(students~gender)/sum(students); tabgen
tabmin<-xtabs(students~minority)/sum(students); tabmin
tabeng<-xtabs(students~native)/sum(students); tabeng
tabten<-xtabs(students~tenure)/sum(students); tabten
tabdiv<-table(division)/sum(table(division)); tabdiv
tabmin<-xtabs(students~minority)/sum(students); tabmin
courses <- length(eval); courses
faculty <- length(unique(prof)); faculty

lower.res<-rbind(wm.course, wsd.course, m.studs, sd.studs, m.peval, female=tabgen[2],
      minority=tabmin[2], 'native english' =tabeng[2], tenure = tabten[2],
      "lower division" = tabdiv[2], courses, faculty); lower.res

colnames(lower.res) <- "lower division"; lower.res

```

****

**Subset Upper**

```{r, warning=FALSE, message=FALSE}

upper <- subset(TeachingRatings, subset=division =="upper")

attach(upper)

wm.course <-weighted.mean(eval,students); wm.course
wsd.course<-my.wsd(students,eval, wm.course, length(eval)) ; wsd.course
m.studs <- mean(allstudents); sd.studs <-sd(allstudents); m.studs
m.peval <-mean(p.eval); m.peval

tabgen<-xtabs(students~gender)/sum(students); tabgen
tabmin<-xtabs(students~minority)/sum(students); tabmin
tabeng<-xtabs(students~native)/sum(students); tabeng
tabten<-xtabs(students~tenure)/sum(students); tabten
tabdiv<-table(division)/sum(table(division)); tabdiv
tabmin<-xtabs(students~minority)/sum(students); tabmin
courses <- length(eval); courses
faculty <- length(unique(prof)); faculty

upper.res<-rbind(wm.course, wsd.course, m.studs, sd.studs, m.peval, female=tabgen[2],
      minority=tabmin[2], 'native english' =tabeng[2], tenure = tabten[2],
      "lower division" = tabdiv[2], courses, faculty);upper.res

colnames(upper.res) <- "upper division" ; upper.res

```


****

**Final Table**

```{r, results='asis'}
tab <-xtable(cbind(all.res, lower.res, upper.res), digits=3)
print(tab, type="html")
```

****

- 

