
Chapter 4: Serving Tables
========================================================


```{r, eval=FALSE}
---
title: "Chapter 4 Brazil table"
author: Murtaza Haider
date: June 20, 2015
---
```

##Using PEW 2010 data set from Global Attitudes##

**Subsetting to Brazil**


**Downloading** data from the website. The download command downloads data in the working directory.

Data are downloaded, loaded, and names of fields are read from the file:

```{r}
setInternet2(TRUE) # solution for https files
download.file("https://sites.google.com/site/econometriks/docs/brazil.RData", "brazil2.RData")
load("brazil2.Rdata")
#load("F:/Research/All_else_being_equal/Chapters/Ch_4_serving tables/R files/Brazil/brazil.RData")
names(brazil)
```

There are 20 variables in the file. 

## Story Line

Never seen more serene vistas than Rio. A city in continuous celebration.
Growing market, identified by Goldman Sachs in BRICS.

*Focus: Internet and social media.*

Data: Brazil

Question: What has age, sex, education, income, ethnicity, and political affiliation to do with the use of Internet and social media?

Illustration of tables
Data set has been cleaned for the following:
* Subsetted
* Column names changed to more common sense explanations 
* new derived variables.

Start with summary statistics. Display the actual output and then show it should be printed. Using Excel and HTML and the Word files.

### Descriptive statistics


Starting with the first variable: country.

*R specific quirk with factors*

Fixing the country variable with refactoring.
```{r}
library(xtable)
library(htmlTable)

```

Uncorrected results
```{r xtable1, results="asis"}
tab<-xtable(cbind("country"=table(brazil$country)))
print(tab, type="html")
```

Corrected results

```{r xtable2, results="asis"}
brazil$country<-factor(brazil$country)
tab<-xtable(cbind("country"=table(brazil$country)))
print(tab, type="html")
```


### What is the status of Internet and social media

Basic descriptives

```{r internet1}
table(brazil$internet)

```

**With xtables**

***Table 4.2***
```{r}

tab<-cbind("internet"=table(brazil$internet))
htmlTable(tab, rowlabel = "User status")


```

```{r internet2, results="asis"}
tab<-xtable(cbind("internet"=table(brazil$internet)))
print(tab, type="html")
```

**Adding more columns**


```{r digital, results="asis"}
tab<-xtable(cbind("internet"=table(brazil$internet),
                  "email"=table(brazil$email),
                  "social media"=table(brazil$social),
                  "cell"=table(brazil$cell)
                  ))
print(tab, type="html")
```

```{r}
library(htmlTable)
tab<-xtable(cbind("internet"=table(brazil$internet),
                  "email"=table(brazil$email),
                  "social media"=table(brazil$social),
                  "cell"=table(brazil$cell)
                  ))
htmlTable(tab, 
          cgroup = c("internet & digital media users"), 
          n.cgroup = c(4))

```




**Using sapply and function**

***Table 4.3***

```{r diginew, results="asis"}
tab<-sapply(brazil[,3:6],function(x) table(x))
htmlTable(tab, cgroup = c("internet & digital media users"), 
          n.cgroup = c(4), rowlabel = "User status")
```


**Proportions**
..

***Table 4.4***
```{r digital2, results="asis"}
tab<-(cbind("internet"=table(brazil$internet),
                  "email"=table(brazil$email),
                  "social media"=table(brazil$social),
                  "cell"=table(brazil$cell)
                  ))

tab1<- addmargins((prop.table(tab,2)*100),1)
htmlTable(txtRound(tab1,1), cgroup = c("internet & digital media users"), 
          n.cgroup = c(4), rowlabel = "User status")
```


**Proportions with sapply, and rounding the output**

***Table 4.5***
```{r propsapply, results="asis"}
tab<-sapply(brazil[,3:6],function(x) table(x))
tab1<- addmargins((prop.table(tab,2)*100),1)
htmlTable(txtRound(tab1,0), cgroup = c("internet & digital media users"), 
          n.cgroup = c(4), rowlabel = "User status")

```


**know your data**
..
*The question for social media is asked of those with internet*

```{r social1, results = "asis"}
tab1<-xtable(table(brazil$social, brazil$internet))
print(tab1, type="html")
tab2<-xtable(table(brazil$cell, brazil$internet))
print(tab2, type="html")
```

```{r}
tab1<-xtable(table(brazil$social, brazil$internet))
htmlTable(tab1, 
          cgroup = c("internet users"), n.cgroup = c(4),
          rgroup=c("social media users"), n.rgroup=c(4))

```

```{r}
tab1<-xtable(table(brazil$social, brazil$internet))
htmlTable(tab1, 
          cgroup = c("internet users"), n.cgroup = c(4),
         rowlabel = "social media users")

```


**Cleaned data set without don't knows and refused**

***Table 4.6***
```{r digital3, results="asis"}
tab<-(cbind("internet"=table(brazil$new.internet),
                  "email"=table(brazil$new.email),
                  "social media"=table(brazil$new.social),
                  "cell"=table(brazil$new.cell)
                  ))
tab1<- addmargins((prop.table(tab,2)*100),1)
htmlTable(txtRound(tab1,1), cgroup = c("internet & digital media users"), 
          n.cgroup = c(4), rowlabel = "User status")
```

**Using the attach option**

```{r attach}
attach(brazil)
```

```{r digital3a, results="asis"}
tab1<-(cbind("internet"=xtabs( ~ new.internet),
                  "email"=xtabs(~new.email),
                  "social media"=xtabs(~new.social),
                  "cell"=xtabs(~new.cell)
                  ))
tab2<-xtable(tab1)
print(tab2, type="html")
tab3<- xtable(addmargins((prop.table(tab1,2)*100),1))
print(tab3, type="html")
```



### Weighted data set using xtabs
```{r digital4, results="asis"}
tab1<-(cbind("internet"=xtabs(weight ~ new.internet),
                  "email"=xtabs(weight~new.email),
                  "social media"=xtabs(weight~new.social),
                  "cell"=xtabs(weight~new.cell)
                  ))
tab2<-xtable(tab1)
print(tab2, type="html")
tab3<- xtable(addmargins((prop.table(tab1,2)*100),1))
print(tab3, type="html")
```



**See the new command with digits=0 in the xtable command**

***Table 4.7***
```{r digital5, results="asis"}
tab1<-(cbind("internet"=xtabs(weight ~ new.internet),
                  "email"=xtabs(weight~new.email),
                  "social media"=xtabs(weight~new.social),
                  "cell"=xtabs(weight~new.cell)
                  ))
htmlTable(txtRound(tab1,0), cgroup = c("internet & digital media users"), 
          n.cgroup = c(4), rowlabel = "User status")

# tab3<- xtable(addmargins((prop.table(tab1,2)*100),1), digits=0)
# print(tab3, type="html")
```


**With SAPPLY and weighted data set**


***Table 4.8***

```{r weightedsapply, results="asis"}
tab1<-sapply(brazil[,14:17],function(x,y) xtabs(y ~ x), y=brazil[,13])
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")

tab3<- txtRound(addmargins((prop.table(tab1,2)*100),1),0)
htmlTable(tab3, rowlabel = "User status")

```

Cross Tabulations
------

Using the recoded variables for income, gender, education and others.

First the tabulating explanatory variables:

**Gender**

```{r gender, results="asis"}
tab1<-cbind("Gender"=xtabs(weight ~ gender))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")
```

```{r gender2, results="asis"}
tab1<-cbind("Gender"=xtabs(weight ~ gender+new.internet))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")

tab3<-xtable(addmargins((prop.table(tab1,1)*100),2),digits=1)
print(tab3, type="html")

```

***Table 4.9 and 4.10***
```{r}
tab1<-cbind(xtabs(weight ~ gender+new.internet))
tab2<-txtRound(tab1, 0)
htmlTable(tab2, 
          cgroup = c("internet users"), n.cgroup = c(2),
          rgroup=c("Gender"), n.rgroup=c(2))

tab3<-addmargins((prop.table(tab1,1)*100),2)
tab4<-txtRound(tab3,1)
htmlTable(tab4, 
          cgroup = c("internet users"), n.cgroup = c(3),
          rowlabel = "Gender")

```


###Education###


```{r edu1, results="asis"}
tab1<-cbind("Education"=xtabs(weight ~ education))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")
```


```{r edu2, results="asis"}
tab1<-cbind("Education"=xtabs(weight ~ education+new.internet))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")
```

***Table 4.11***
```{r}
tab1<-cbind("Education"=xtabs(weight ~ education+new.internet))
tab2<-txtRound(addmargins((prop.table(tab1,1)*100),2),1)
htmlTable(tab2, cgroup = c("internet users"), n.cgroup = c(3),
          rowlabel = "Education attainment")
```


**Income**

```{r inc1, results="asis"}
tab1<-cbind("income"=xtabs(weight ~ income))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")
```

```{r inc2, results="asis"}
tab1<-cbind("income"=xtabs(weight ~ income+new.internet))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")

tab3<-xtable(addmargins((prop.table(tab1,1)*100),2),digits=1)
print(tab3, type="html")

```

***Table 4.12***

```{r}

tab1<-cbind("Education"=xtabs(weight ~ income+new.internet))
tab2<-txtRound(addmargins((prop.table(tab1,1)*100),2),1)
htmlTable(tab2, cgroup = c("internet users"), n.cgroup = c(3),
          rowlabel = "Income level")

```

**Age**

```{r age1, results="asis"}
tab1<-cbind("age"=xtabs(weight ~ new.age))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")
```

```{r age2, results="asis"}
tab1<-cbind("age"=xtabs(weight ~ new.age+new.internet))
tab2<-xtable(tab1, digits=0)
print(tab2, type="html")

tab3<-xtable(addmargins((prop.table(tab1,1)*100),2),digits=1)
print(tab3, type="html")

```

***Table 4.13***

```{r}
tab1<-cbind("Age cohort"=xtabs(weight ~ new.age+new.internet))
tab2<-txtRound(addmargins((prop.table(tab1,1)*100),2),1)
htmlTable(tab2, cgroup = c("internet users"), n.cgroup = c(3),
          rowlabel = "Age cohorts")

```




**Age + Gender**

```{r age_gender}
round(ftable(xtabs(weight~gender+new.age+new.internet)),0)
```

## Merging it all in a Binary Logit model

**GLM.1 --> unweighted model GLM.2 --> weighted model**

```{r, results='hide' }
GLM.1 <- glm(new.internet ~ gender + new.age + new.cell + new.inc, family=binomial(logit), data=brazil)
GLM.2 <- glm(new.internet ~ gender + new.age + new.cell + new.inc, family=binomial(logit), 
    weights=weight, data=brazil)
```

```{r, results='asis'}

logit1<-xtable(summary(GLM.1))
print(logit1, type="html")

logit2<-xtable(summary(GLM.2))
print(logit2, type="html")

expon<-cbind("unweighted model"=exp(coef(GLM.1)), "weighted model"=exp(coef(GLM.2)))
print(xtable(expon,digits=3), type="html")
```


Stargazer with Brazil data
===



```{r, message=FALSE, echo=FALSE}
library(stargazer)
```

## Merging it all in a Binary Logit model

**GLM.1 --> unweighted model GLM.2 --> weighted model**

```{r, temp1, results='hide' }
GLM.1 <- glm(new.internet ~ gender + new.age + new.cell + new.inc, family=binomial(logit), data=brazil)
GLM.2 <- glm(new.internet ~ gender + new.age + new.cell + new.inc, family=binomial(logit), 
    weights=weight, data=brazil)
```




```{r, temp2, results='asis'}

stargazer(GLM.1, GLM.2, type="html", align=TRUE, no.space=TRUE, 
          dep.var.labels=c("Internet use (unweighted data), Internet use (weighted data)"),
          covariate.labels=c("Female", "seniors","young adults", "youth", "own cell phone", 
                             "low income hhld", "medium income hhld"))
```



